---
title: "Covid91 vaccine study Final2023"
author: "Marta Marriott"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
male = filter(FinalData, sex == "M")
female = filter(FinalData, sex == "F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

#Dataset details

First,the varibales from the FinalData had to be changed from character to factor variable using the  dplyr code:

FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))

I also used the Cat_Cat.rmd help file accomplish this.

Next, the  subgroups had to be made into borader categories in order to get the proper results. Male and Female were made into subgroups, as well as gay/LBGTQ and those to answred yes to using drugs. The categories were made in order to see whether or not these groups should or should not take the new covid-19 vaccine.

In order to create an effectiveness interval for all four categories each group was made into seperate variables.

## Males

### Barplot

```{r}
dd2 <- male %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
The plot represents the male population and the relationship of that poulation contracting covid-19 after getting the vaccine or placebeo. 

The first plot show how many males were given the placebo and the vaccine. Only small portion who took the vaccine or placebo caught covid, while nearly 10,000 of those to received the vaccine or placebo did not catch covid-19. The second graph shows that nearly 80 percent that took the placebo got infected, while only 30 percent who took the vaccine became infected. Nearly 50 percent in each group did not become infected.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=male)
rowPerc(table1)
colPerc(table1)
```

The numerical results show the percentages of males who contracted covid-19 after taking the vaccine or placebo. Nearly 70 percent who took the placebo contracted covid-19 and 30 percent took the vaccine caught covid. However, only 50 percent did not contract covid-19 after the placebo and roughly 50 percent did not contract covid-19 after the vaccine. Because there was a higher instance of contracting covid-19 when the the placebo, it shows evidence that the vaccine is safe and effective for males.

### Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
malesfish <- fisher.test(table1)
malesfish
```

The inferential results show the infection rate of those who contracted covid-19 after taking the vaccine was low, while those who took the placebo had a higher contraction rate. the effectiveness confidence internal was greater than 1 according to the null hypthesis, so there is evidence of effectiveness of the 
vaccine among males.

Therefor, the graph, numerical and inferential analysis, show that the vaccine does offer some effectiveness among males ans is therefore safe to take.

## Females



## LGBTQ



## Druggies


# Overall Results and Conclusions