---
title: "Covid91 vaccine study Final2023"
author: "Marta Marriott"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
male = filter(FinalData, sex == "M")
female = filter(FinalData, sex == "F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

#Dataset details

First,the varibales from the FinalData had to be changed from character to factor variable using the  dplyr code:

FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))

I also used the Cat_Cat.rmd help file accomplish this.

Next, the  subgroups had to be made into borader categories in order to get the proper results. Male and Female were made into subgroups, as well as gay/LBGTQ and those to answred yes to using drugs. The categories were made in order to see whether or not these groups should or should not take the new covid-19 vaccine.

In order to create an effectiveness interval for all four categories each group was made into seperate variables.

## Males

### Barplot

```{r}
dd2 <- male %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
The plot represents the male population and the relationship of that poulation contracting covid-19 after getting the vaccine or placebeo. 

The first plot show how many males were given the placebo and the vaccine. Only small portion who took the vaccine or placebo caught covid, while nearly 10,000 of those to received the vaccine or placebo did not catch covid-19. The second graph shows that nearly 80 percent that took the placebo got infected, while only 30 percent who took the vaccine became infected. Nearly 50 percent in each group did not become infected.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=male)
rowPerc(table1)
colPerc(table1)
```

The numerical results show the percentages of males who contracted covid-19 after taking the vaccine or placebo. Nearly 70 percent who took the placebo contracted covid-19 and 30 percent took the vaccine caught covid. However, only 50 percent did not contract covid-19 after the placebo and roughly 50 percent did not contract covid-19 after the vaccine. Because there was a higher instance of contracting covid-19 when the the placebo, it shows evidence that the vaccine is safe and effective for males.

### Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
malesfish <- fisher.test(table1)
malesfish
```

The inferential results show the infection rate of those who contracted covid-19 after taking the vaccine was low, while those who took the placebo had a higher contraction rate. the effectiveness confidence internal was greater than 1 according to the null hypthesis, so there is evidence of effectiveness of the 
vaccine among males.

Therefor, the graph, numerical and inferential analysis, show that the vaccine does offer some effectiveness among males ans is therefore safe to take.

## Females

### Barplot

```{r}
dd2 <- female %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The plot represents the female population and the relationship of that poulation contracting covid-19 after getting the vaccine or placebeo. 

The first plot show how many males were given the placebo and the vaccine. Only small portion who took the vaccine or placebo caught covid, while nearly 10,000 of those to received the vaccine or placebo did not catch covid-19. The second graph shows that nearly 60 percent that took the placebo got infected, while only 40 percent who took the vaccine became infected. Nearly 50 percent in each group did not become infected.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=female)
rowPerc(table1)
colPerc(table1)
```

The numerical resluts for females show similar results to that of the male population. The results show a high infection rate for those that took the placebo than those who took the vaccine. In fcat, only 40 percent who took the vaccine contracted covid. Therefore, the drug is effect for females to take.

### Inferntial results

```{r}
chisq.test(table1)
chisqtestGC(table1)
femalesfish <- fisher.test(table1)
femalesfish
```

The inferential results coincide with the numerical results. The female population who took the vaccine, had a lower contraction rate that those who took the placebo.

Based on the results of the barplot, numerical and inferential results, the vaccine seems to be safe and effective for the female opoulation.

## LGBTQ

### Barplot

```{r}
dd2 <- gay %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barplot shows a shift in the effectiveness in contraction rate among those in the LBGTQ community. The contraction rate of covid-19 of those who  identify ad LGBTQ or gay is over 80 percent, compared to the previous two groups. The placebo and vaccinated groups who do not contract the illness, seem to be much lower at 50 percent.

# Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

The numerical data correlates with the barplot, that the contraction rate for those who received that vaccine and still contracted the illness is relatively high, thus giving inclination that that vaccine for the LGBTQ community may be unsafe or less effective.

### Inferential results


```{r}
chisq.test(table1)
chisqtestGC(table1)
gaysfish <- fisher.test(table1)
gaysfish
```

The data from the inferential results seem to reflect that of the barplot and the numerical results. Among the LGBTQ community, there seems to be a lower effectiveness rate and may be considered unsafe.

The barplot, numerical, and inferential results, seem to show that the vaccine may be slightly less effective.

## Druggies

### Barplot

```{r}
dd2 <-  Druguser %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barplot seems to show that the group who identify as drug users, also have a higher instance of contracting covid-19  when vaccinated. This group also has an 80 percent contraction rate among the vaccinated group and less than 50 percent than those who are not vaccinated.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=Druguser)
rowPerc(table1)
colPerc(table1)
```

The numerical results seem to reflect that of the bargraph. Those who identify as drug users have a higher contraction rate than those who do not identify as drug users or are not vaccinated.

### Inferntial results

```{r}
chisq.test(table1)
chisqtestGC(table1)
Drugusersfish <- fisher.test(table1)
Drugusersfish
```

The inferential results show that that those who received the placebo had a lower contraction rate than those who received the vaccine. Therefore, we can conclude that the vaccine for drug users may be ineffective.

Based on the date results for the drug user group, the vaccine may be unsafe and ineffective.

# Overall Results and Conclusions